---
title: "2.2 Graphical summaries of data"
author: "Luis Talavera"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
```

## Exercises
**2.10 Generating Electricity** In 2012 in the United States,
most electricity was generated from coal (37%), natural
gas (30%), or nuclear power plants (19%). Hydropower
accounted for 7% of the total electricity produced; other
renewable sources such as wind or solar power accounted
for 5%. Other nonrenewable sources (such as petroleum)
made up the remaining 2%. (Source: http://www.eia.gov/
electricity/annual/html/epa_01_01.html)  
  a. Display this information in a bar graph.  
  b. Which is easier to sketch relatively accurately, a pie
chart or a bar chart?  
  c. What is the advantage of using a graph to summarize
the results instead of merely stating the percentages for
each source?  
  d. What is the modal category?  
  
#### Ans.
```{r}
energy_sources <- tibble(energy_source = c("coal", "natural gas", "hydropower",
                                           "other_renewable", "nonrenewable"),
                         elec_generated_percent = c(37,30,19,7,5))

g <- ggplot(energy_sources, aes(x = reorder(energy_source, -elec_generated_percent), y = elec_generated_percent))
g <- g + geom_bar(aes(fill = energy_source), stat="identity") +
     labs(title="Electricity generated sources",
        x ="Energy source", y = "Electricity percentage") +
     theme(legend.position = "none")
g
```

