---
title: "2.2 Graphical summaries of data"
author: "Luis Talavera"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(ggpubr)
```

## Exercises
**2.10 Generating Electricity** In 2012 in the United States,
most electricity was generated from coal (37%), natural
gas (30%), or nuclear power plants (19%). Hydropower
accounted for 7% of the total electricity produced; other
renewable sources such as wind or solar power accounted
for 5%. Other nonrenewable sources (such as petroleum)
made up the remaining 2%. (Source: http://www.eia.gov/
electricity/annual/html/epa_01_01.html)  
  a. Display this information in a bar graph.  
  b. Which is easier to sketch relatively accurately, a pie
chart or a bar chart?  
  c. What is the advantage of using a graph to summarize
the results instead of merely stating the percentages for
each source?  
  d. What is the modal category?  
  
#### Ans.
  a.  
```{r, out.width="75%", fig.align = 'center'}
energy_sources <- tibble(energy_source = c("coal", "natural gas", "hydropower",
                                           "other_renewable", "nonrenewable"),
                         elec_generated_percent = c(37,30,19,7,5))

g <- ggplot(energy_sources, aes(x = reorder(energy_source, -elec_generated_percent), y = elec_generated_percent))
g <- g + geom_bar(aes(fill = energy_source), stat="identity") +
     labs(title="Electricity generated sources",
        x ="Energy source", y = "Electricity percentage") +
     theme(legend.position = "none")
g
```
  b. A bar chart
  c. The modal category can be easily identified as well as the distribution of the data.
  d. Coal
  
**2.11 What do alligators eat?** The bar chart is from a study investigating the factors that 
  influence alligators’ choice of food. For 219 alligators captured
in four Florida lakes, researchers classified the
primary food choice (in volume) found in the
alligator’s stomach in one of the categories—fish,
invertebrate (snails, insects, crayfish), reptile
(turtles, baby alligators), bird, or other (amphibian, mammal, plants). (Data available on the book’s
website.)  
  a. Is primary food choice categorical or quantitative?  
  b. Which is the modal category for primary food
choice?  
  c. About what percentage of alligators had fish as the
primary food choice?  
  d. This type of bar chart, with categories listed in order of
frequency, has a special name. What is it?  

```{=latex}
\begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{"./grafica_barra1.jpg"}
  \vspace{-0.5cm}
  \caption{Primary Food Choice of Alligators}
  \label{fig:Primary Food Choice of Alligators}
\end{figure}
```
  
#### Ans.
  a. Primary food is a categorical food
  b. Fish
  c. ~42%
  d. Pareto chart
  
**2.12 Weather stations** The pie chart (constructed using
EXCEL) shown portrays the regional distribution of
weather stations in the United States.  
  a. Do the slices of the pie portray (i) variables or
(ii) categories of a variable?  
  b. Identify what the two numbers mean that are shown
for each slice of the pie.  
  c. Without inspecting the numbers, would it be easier to
identify the modal category by using this graph or the
corresponding bar graph? Why?  

```{=latex}
\clearpage
\begin{figure}[hbtp]
  \centering
  \includegraphics[width=\textwidth]{"./grafica_pastel.jpg"}
  \vspace{-0.5cm}
  \caption{Regional Distribution of Weather Stations}
  \label{fig:Regional Distribution of Weather Stations}
\end{figure}
```

#### Ans.
  a. Categories of a variable.
  b. The absolute and relative frequency.
  c. It would be easy using a bar graph, without the numbers both Midwest and West can be considered as the modal category.
  
**2.13 France is most popular holiday spot** Which countries are
most frequently visited by tourists from other countries?
The table shows results according to Travel and Leisure
magazine (2005).    
  a. Is country visited a categorical or a quantitative
variable?  
  b. In creating a bar graph of these data, would it be most
sensible to list the countries alphabetically or in the
form of a Pareto chart? Explain.  
  c. Does either a dot plot or stem-and-leaf plot make
sense for these data? Explain.  

```{=latex}
\begin{figure}[hbtp]
  \centering
  \includegraphics[width=\textwidth]{"./tabla_paises_mas_visitados.jpg"}
  \vspace{-0.5cm}
  \caption{Regional Distribution of Weather Stations}
  \label{fig:Regional Distribution of Weather Stations}
\end{figure}
```

#### Ans.
  a. A categorical variable.
  b. Pareto chart form is better, it displays the most visited countries at the beginning of the table
  c. They does not make sense. These graphs are not for categorical data.
  
**2.14 Pareto chart for fatal shark attacks** The data shown in
Exercise 2.9 give frequencies of fatal shark attacks for
different regions. Using software or sketching, construct
a bar graph, ordering the regions (i) alphabetically and
(ii) as in a Pareto chart. Which do you prefer? Why?
  
```{r, out.width="75%", fig.align = 'center'}
fatal_shark_attacks <- tibble(region = c("Florida", "Hawaii", "California", 
                                    "Australia", "South Africa", "Réunon Island", 
                                    "Brazil", "Bahamas", "Other"),
                         frequency = c(2,2,4,15,13,6,4,6,11))

g <- ggplot(fatal_shark_attacks, aes(x = region, y = frequency))
g <- g + geom_bar(aes(fill = region), stat="identity") +
     labs(title="Fatal shark attacks",
        x ="Region", y = "Frequency") +
     theme(legend.position = "none") +
scale_x_discrete(guide = guide_axis(n.dodge=2))

h <- ggplot(fatal_shark_attacks, aes(x = reorder(region, -frequency), y = frequency))
h <- h + geom_bar(aes(fill = region), stat="identity") +
     labs(title="Fatal shark attacks",
        x ="Region", y = "Frequency") +
     theme(legend.position = "none") +
scale_x_discrete(guide = guide_axis(n.dodge=2))

figures <- ggarrange(g, h,
                    ncol = 2, nrow = 1)
figures
```

#### Ans.  

I prefer Pareto form chart due to displays most frequent categories at the left and least important categories at the right.

**2.15 Sugar dot plot** For the breakfast cereal data given in
Table 2.3, a dot plot for the sugar values (in grams) is
shown:  
  a. Identify the minimum and maximum sugar values.  
  b. Which sugar outcomes occur most frequently? What
are these values called?  

```{=latex}
\begin{figure}[hbtp]
  \centering
  \includegraphics[width=\textwidth]{"./azucar_en_cereales.jpg"}
  \vspace{-0.5cm}
  \caption{Breakfast cereal sugar data}
  \label{fig:Breakfast cereal sugar data}
\end{figure}
```

#### Ans. 
  a. 0g and 18g
  b. 3, 4, 11, 12, 14. These values are called modes.
  
**2.16 Spring break hotel prices** For a trip to Miami, Florida,
over spring break in 2014, the data below (obtained from
travelocity.com) show the price per night (in U.S. dollars)
for various hotel rooms.

```{=latex}
\begin{table}[htbp]
\centering
\begin{tabular}{lllllllll}
239 & 237 & 245 & 310 & 218 & 175 & 330 & 196 & 178 \\
245 & 255 & 190 & 330 & 124 & 162 & 190 & 386 & 145
\end{tabular}
\end{table}
```

  a. Construct a stem-and-leaf plot. Truncate the data to
the first two digits for purposes of constructing the
plot. For example, 239 becomes 23.  
  b. Reconstruct the stem-and-leaf plot in part a by using
split stems; that is, two stems of 1, two stems of 2, and
so on. Compare the two stem-and-leaf plots. Explain
how one may be more informative than the other.  
  c. Sketch a histogram by hand (or use software), using 6
intervals of length 50, starting at 100 and ending at 400.
What does the plot tell about the distribution of hotel
prices? (Mention where most prices tend to fall and
comment about the shape of the distribution.)  

#### Ans.  

  a.  
```{r, fig.align = 'center', comment=NA}
data <- c(239, 237, 245, 310, 218, 175, 330, 196, 178, 
          245, 255, 190, 330, 124, 162, 190, 386, 145)

stem(data)
```

  b.  
```{r, fig.align = 'center', comment=NA}
stem(data, scale=2)
```

  c.  
```{r, out.width="50%", fig.align = 'center', comment=NA}
hist(data, breaks = 6, xlim = c(100,400))
```

**Exercises 2.17, 2.18 and 2.19 were similar, since a leaf-stem chart is rarely used I decided to skip those exercises**  

**2.20 Histogram for sugar** For the breakfast cereal data,
the figure below shows a histogram for the sugar values
in grams.  
  a. Identify the intervals of sugar values used for the plot.  
  b. Describe the shape of the distribution. What do you
think might account for this unusual shape? (Hint:
How else are the cereals classified in Table 2.3?)  
  c. What information can you get from the dot plot or
stem-and-leaf plot of these data shown in Exercises
2.15 and 2.19 that you cannot get from this plot?  
  d. This histogram shows frequencies. If you were to
construct a histogram by using the percentages for
each interval, how (if at all) would the shape of this
histogram change?  

```{=latex}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{"./conteo_azucar.jpg"}
  \vspace{-0.5cm}
  \caption{Breakfast cereal data}
  \label{fig:Breakfast cereal data}
\end{figure}
```

#### Ans.
  a. 0-2, 2-4, 4-6, 6-8, 8-10, 10-12, 12-14, 14-16, 16-18
  b. Is a bimodal shape
  c. The exact numbers
  d. It would only change the y axis but the shape will keep the same  
  
**2.21 Shape of the histogram** For each of the following variables, indicate whether you would expect its histogram to
be symmetric, skewed to the right, or skewed to the left.
Explain why.  
  a. The price of a certain model of smartwatch in different
stores in your district  
  b. The amount of time students use to take an exam in
your school  
  c. The grade point average (GPA) in your academic
program this year  
  d. The salary of all the employees in a company.  
  
#### Ans.
  a. symmetric, the majority of stores will sell the smartwatch at relatively close prices.
  b. Skewed to the left, most of students spend the same amount of time to study, just a few who didn't compromised study for a short time.
  c. symmetric, most of students will share GPA.
  d. Skewed to the right, the differences betwwen salaries could be so large.
  
**2.22 More shapes of histograms** Repeat the preceding
exercise for  
  a. The winner’s score in a basketball game during an
NBA season.  
  b. The distance from home to school for students in a
specific school  
  c. The number of attempts a young adult needs to pass a
driving license test  
  d. The number of times an individual requests a password
reset for the forgotten password of his or her email
account.  

#### Ans.
  a. symmetric
  b. symmetric
  c. Skewed to the right.
  d. Skewed to the right.
  
**2.23 Gestational Period** The Animals data set at the
book’s website contains data on the length of the gestational period (in days) of 21 animals. (Source: Wildlife
Conservation Society)  
  a. Using software, plot a histogram of the gestational
period.  
  b. Do you see any observation that is unusual? Which
animal is it?  
  c. Is the distribution right-skewed or left-skewed?  
  d. Try to override the default setting and plot a histogram with only very few (e.g., 3 or 4) intervals and
one with many (e.g., 30) intervals. Would you prefer
either one to the histogram created in part a? Why or
why not?  

#### Ans.

a.    
```{r, message=FALSE, out.width="75%", fig.align = 'center'}
animals <- read_csv("../datasets/Animals.csv")

ggplot() + aes(animals$gestation)+ geom_histogram(colour="black", fill="white")
```
  b. Ass, is humans
  c. Right-skewed
  d. 
  
```{r, out.width="75%", fig.align = 'center'}
gestation_days <- animals$gestation

p <- ggplot() + aes(gestation_days)+ geom_histogram(colour="black", fill="blue", bins = 5)
q <- ggplot() + aes(gestation_days)+ geom_histogram(colour="black", fill="red", bins = 50)

figures <- ggarrange(p, q,
                    ncol = 2, nrow = 1)
figures
```
  I think is case-dependent, more bins more smooth and more informative but too many bins make the plot difficult to understand.

**2.24 How often do students read the newspaper?** Question
14 on the class survey (Activity 3 in Chapter 1) asked,
“Estimate the number of times a week, on average, that
you read a daily newspaper.”  
  a. Is this variable continuous or discrete? Explain.  
  b. The histogram shown gives results of this variable when
this survey was administered to a class of 36 University
of Georgia students. Report the (i) minimum response,
(ii) maximum response, (iii) number of students who
did not read the newspaper at all, and (iv) mode.  
  c. Describe the shape of the distribution.  
  
  
```{=latex}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.75\textwidth]{"./frequencia_periodico.jpg"}
  \vspace{-0.5cm}
  \caption{Number Times/Week Read Newspaper 36 Gerogia students}
  \label{fig:Number Times/Week Read Newspaper 36 Gerogia students}
\end{figure}
```
  
#### Ans.
  a. Discrete, is a countable quantity (1, 2, 3, 4, etc).
  b. 
    i. 0.  
    ii. 9.  
    iii. 2.  
    iv. 3  
  c.Skewed to the right
  
**2.25 Blossom widths** A data set (available at the book’s
website) analyzed by the famous statistician R. A. Fisher
consisted of measurements of different varieties of iris
blossoms. Histograms representing the widths of the petals of the two species, Iris setosa and Iris versicolor, are
on the next page.

```{=latex}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.75\textwidth]{"./ancho_petalos.jpg"}
  \vspace{-0.5cm}
  \caption{Iris setosa and Iris veriscolor widths}
  \label{fig:Iris setosa and Iris veriscolor widths}
\end{figure}
```

  a. Describe the shape of the distribution of setosa petal
widths. 
  b. Describe the shape of the distribution of versicolor
petal widths.
  c. Of the 50 versicolor blossoms in the data set, approximately what percentage has a petal width less than
3.2 inches?
  d. Is it possible to determine accurately the percentage of
versicolor blossoms with a width of more than 3 inches?
Why or why not?  

#### Ans.  

  a. Setosa: Skewed to the left. 
  b. Versicolor: Skewed to the right.
  c. ~ 92%
  d. Using the graph, yes it shows the frequency of blossoms with a width of more than 3 inches is almost 5.
  
  
**2.26 Central Park temperatures** The first figure shows a
histogram of the Central Park, New York, annual average
temperatures from 1869 to 2012.

```{=latex}
\clearpage
\begin{figure}[htpb!]
  \centering
  \includegraphics[width=0.75\textwidth]{"./temperatura_anual.jpg"}
  \vspace{-0.5cm}
  \caption{Central Park, New York, annual average
temperatures from 1869 to 2012.}
  \label{fig:Central Park, New York, annual average
temperatures from 1869 to 2012.}
\end{figure}
```

  a. Describe the shape of the distribution.
  b. What information can the time plot above show that a
histogram cannot provide? 
  c. What information does the histogram show that a time
plot does not provide?

#### Ans.
  a. Symmetric
  b. A decreasing or increasing tendency of averages.
  c. Where most of the data falls into.
  
**2.27 Is whooping cough close to being eradicated?** In the
first half of the 20th century, whooping cough was a frequently occurring bacterial infection that often resulted in
death, especially among young children. A vaccination for
whooping cough was developed in the 1940s. How effective
has the vaccination been in eradicating whooping cough?
One measure to consider is the incidence rate (number of
infected individuals per 100,000 population) in the United
States. The table shows incidence rates from 1925 to 1970.

```{=latex}
\begin{figure}[htpb!]
  \centering
  \includegraphics[width=0.75\textwidth]{"./whooping.jpg"}
  \vspace{-0.5cm}
  \caption{Incidence Rates for Whooping Cough, 1925–1970}
  \label{fig:Incidence Rates for Whooping Cough, 1925–1970}
\end{figure}
```

  a. Sketch a time plot. What type of trend do you observe?
Based on the trend from 1945 to 1970, was the whooping cough vaccination proving effective in reducing the
incidence of whooping cough?
  b. The figure shown is a time plot of the incidence rate from
1980 to 2012, based on numbers reported by the Center for
Disease Control and Prevention.3 Describe the time series.
Is the United States close to eradicating whooping cough?
What do you think contributes to the recent trend? (Data
available on the book’s website in the pertussis.csv file.)

```{=latex}
\begin{figure}[htpb!]
  \centering
  \includegraphics[width=0.75\textwidth]{"./whooping_1980.jpg"}
  \vspace{-0.5cm}
  \caption{Incidence Rates for Whooping Cough, 1980-2012}
  \label{fig:Incidence Rates for Whooping Cough, 1980-2012}
\end{figure}
```

  c. Would a histogram of the incidence rates since 1935
address the question about the success of the vaccination for whooping cough? Why or why not?

#### Ans.
  a.
```{r, out.width="75%", fig.align = 'center'}
incidence_rates <- tibble(year = seq(1925, 1970, 5), 
                          rates = c(131.2,135.6,141.9,139.6,101.0,80.1,38.2,8.3,3.5,2.1))

ggplot(data = incidence_rates, aes(x = year, y = rates)) +
  geom_line(color = "#00AFBB", size = 1) +
  labs(x="Year", y="Rate per 100,000")
```
  Yes, the rates has been decreasing over the years
  
  b. No, US is not close to eradicate the disease due to people stop from getting the vaccine.
  c. By the plots, we could se the campaign was effective and after the campaign stops the rates increased.
  
**2.28 Warming in Newnan, Georgia?** Access the Newnan, GA
Temps file on the book’s website, which reports the average
annual temperatures during the 20th century for Newnan,
Georgia. Construct a time plot to investigate a possible
trend over time. Is there evidence of climate change?

#### Ans.

```{r, out.width="60%", fig.align = 'center', message=FALSE}
newnan_ga_temps <- read_csv("../datasets/newnan_ga_temps.csv")

ggplot(data = newnan_ga_temps, aes(x = Year, y = Temp)) +
  geom_line(color = "#00AFBB", size = 1) +
  labs(x="Year", y="Temperature")
```

Well, the temperature in Newman Georgia has been decreasing over the years, but a temperature records from an specific place can't be used to determine a global climate change.