---
title: "2.3 Measuring the Center of Quantitative Data"
author: "Luis Talavera"
date: "`r Sys.Date()`"
output:
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(ggpubr)
```

## Exercises

**2.29 Median versus mean** For each of the following variables, would
you use the median or mean for describing the center of the
distribution? Why? (Think about the likely shape of the distribution.)\
a. Salary of employees of a university\
b. Time spent on a difficult exam\
c. Scores on a standardized test

#### Ans.

a.  Median (right-skewed)
b.  Median (left-skewed)
c.  Mean (symmetric)

**2.30 More median versus mean** For each of the following variables,
would you use the median or mean for describing the center of the
distribution? Why? (Think about the likely shape of the distribution.)\
a. Amount of liquid in bottles of capacity one liter b. The salary of
all the employees in a company\
c. Number of requests to reset passwords for individual email accounts.

#### Ans.

a.  Mean (symmetric)
b.  Median (right-skewed)
c.  Median (right-skewed)

**2.31 More on CO2 emissions** The Energy Information Agency reported
the CO2 emissions (measured in gigatons, Gt) from fossil fuel combustion
for the top 10 emitting countries in 2011. These are China (8 Gt), the
United States (5.3 Gt), India (1.8 Gt), Russia (1.7 Gt), Japan (1.2 Gt),
Germany (0.8 Gt), Korea (0.6 Gt), Canada(0.5 Gt), Iran (0.4 Gt), and
Saudi Arabia (0.4 Gt).\
a. Find the mean and median CO2 emission.\
b. The totals reported here do not take into account a nation's
population size. Explain why it may be more sensible to analyze per
capita values, as was done in Example 11.

#### Ans.

a.  

```{r, comment=""}
co2_emissions <- c(8,5.3,1.8,1.7,1.2,0.7,0.6,0.5,0.4,0.4)
m <- mean(co2_emissions)
md <- median(co2_emissions)

message(paste("Mean: ", m))
message(paste("Median: ", md))
```

b.  Nations with a larger population will have larger co2 emissions,
    analyzing per capita values will give us another point of view more
    accurate.

**2.32 Resistance to an outlier** Consider the following three sets of
observations: Set 1: 8, 9, 10, 11, 12 Set 2: 8, 9, 10, 11, 100 Set 3: 8,
9, 10, 11, 1000\
a. Find the median for each data set. b. Find the mean for each data
set. c. What do these data sets illustrate about the resistance of the
median and mean?

#### Ans.

a.  

```{r, comment=""}
s1 <- c(8, 9, 10, 11, 12)
s2 <- c(8, 9, 10, 11, 100)
s3 <- c(8, 9, 10, 11, 1000)

message(paste("Median set 1: ", median(s1)))
message(paste("Median set 2: ", median(s2)))
message(paste("Median set 3: ", median(s3)))
```

b.  

```{r, comment=""}
message(paste("Median set 1: ", mean(s1)))
message(paste("Median set 2: ", mean(s2)))
message(paste("Median set 3: ", mean(s3)))
```

c.  The median is resistance to outliers but the mean not.

**2.33 Weekly earnings and gender** In New Zealand, the mean and median
weekly earnings for males in 2009 was \$993 and \$870, respectively and
for females, the mean and median weekly earnings were \$683 and \$625,
respectively (www.nzdotstat.stats.govt.nz). Does this suggest that the
distribution of weekly earnings for males is symmetric, skewed to the
right, or skewed to the left? What about the distribution of weekly
earnings for females? Explain.

#### Ans.

If the mean and median are equal the distribution is symmetric, if the
mean is below the median the distribution is left-skewed and if the mean
is above the median the distribution is right-skewed. Therefore the
distribution of weekly earning for males is right-skewed and for females
is the same.

**2.34 Labor dispute** The workers and the management of a company are
having a labor dispute. Explain why the workers might use the median
income of all the employees to justify a raise but management might use
the mean income to argue that a raise is not needed.

#### Ans.

A company salaries distribution is right-skewed, indicating that most of
the workers earn almost the same and a few ones earn more by an
incredible difference. The median is not sensitive to outliers (the
salary of management and chiefs) so the workers will use the median
justify a raise since the median will reflect their current income and
management will use mean to hide this difference between salaries.

### Note

The rest of the exercises are almost the same, calculating median and
mean, comparing, drawing conclusions, the only differences were that
they gave you the information or you had to obtain it, since it is a
subject that I understood quickly I will choose to skip those remaining
exercises.
